Hydras, Ordinals \& Co. \\ A library in Coq of entertaining formal mathematics

### 2021-10-06 doc/JFLA2022/paper.tex
Changing some  figure positions

 Prior stages of this project have already been presented at
 JFLA~\cite{PCiota, JFLA2018paper}.

We present recent evolutions of the library: new results, interaction within the \community project~\cite{CoqCommunity}, and documentation generated with \alectr~\cite{alectryonpaper, alectryongithub}.

\section{Changes in the library}
The 2018 article~\cite{JFLA2018paper} contains a formal proof of  a variant of Theorem~\ref{kp:thm2}:

@@

Considering measures applicable to \emph{all} battles allowed us to
focus on battles where the hydra can make an arbitrary number of copies at any time, which made our proof by contradiction artificially easier.
Unfortunately, the examples  most commonly shown in the litterature
(see for instance~\cite{KP82, bauer2008, BauerHydra}) 
assume that the hydra grows $n$ copies at step $n$ of the game, which is incompatible with our proof.

\vspace{6pt}

We prove now that Theorem~\ref{thm3} still holds with these typical battles by borrowing new combinatorial results from~\cite{KS81}.
Without loss of generality, we assume the following restrictions:

@@

Using $H'$s equations as rewrite rules, we can study a realistic example. We take the hydra of figure~\ref{fig:start} and $i=0$ as initial configuration. 
By a sequence of rewritings and inductions, we prove that the number of steps of the considered battle is greater or equal than $2^{2^N}$, where $N=2^{70}-1$.

More generally, we prove that, for $\alpha\geq\omega^\omega$
the function computing the length of the battle starting with the configuration $(\alpha,i)$ is not primitive recursive.

@@

The \Hydras book is written in LaTeX, but it makes very frequent references (about once per page, 274 snippets over 281 pages) to parts of the \coq development, showing definitions (Fig.~\ref{fig:Hprime}), lemmas (Fig.~\ref{fig:Ex42E0}), proof scripts (Fig.~\ref{fig:Ex42-schutte}), and computation results (Fig.~\ref{fig:Pow-evalPow17LetIn}). The order in which these references appear in the book is independent of the structure of our libraries. Thus, we chose not to include our textbook as comments in the libraries (solution \emph{à la} Coqdoc).
%\TODO[Pierre]{is the preceding sentence OK?}

@@

@@

\begin{figure}
  \centering
  \inputsnippets{Pow/evalPow17LetIn}
  \caption{\label{fig:Pow-evalPow17LetIn}Automatically capturing the output of computations}
\end{figure}

@@

The build artifacts are stored in the \community binary cache on Cachix~\cite{cachix}.
%
This means that a given version never has to be built twice and Continuous Integration can be almost instantaneous when no change has been made (e.g., after merging a pull request).

>>>>

 Prior stages of this project have already been presented at
 JFLA~\cite{PCiota, JFLA2018paper}.
We present recent evolutions of the library: new results, interaction within the \community project~\cite{CoqCommunity}, and documentation generated with \alectr~\cite{alectryonpaper, alectryongithub}.

\section{Changes in the library}
The 2018 article~\cite{JFLA2018paper} presented a formal proof of  a variant of Theorem~\ref{kp:thm2}:

@@

Considering measures applicable to \emph{all} battles allowed us to
focus on battles where the hydra can make an arbitrary number of copies at any time, which made our proof by contradiction artificially easier.
Unfortunately, the examples  most commonly shown in the litterature
(see for instance~\cite{KP82, bauer2008, BauerHydra}) 
assume that the hydra grows $n$ copies at step $n$ of the game, which is incompatible with our proof.
We prove now that Theorem~\ref{thm3} still holds with these typical battles by borrowing new combinatorial results from~\cite{KS81}.
Without loss of generality, we assume the following restrictions:

@@

Using $H'$s equations as rewrite rules, we can study a realistic example. We take the hydra of figure~\ref{fig:start} and $i=0$ as initial configuration. 
By a sequence of rewritings and inductions, we prove that the number of steps of the considered battle is greater or equal than $2^{2^N}$, where $N=2^{70}-1$.
More generally, we prove that, for $\alpha\geq\omega^\omega$
the function computing the length of the battle starting with the configuration $(\alpha,i)$ is not primitive recursive.

@@

The \Hydras book is written in LaTeX, but it makes very frequent references (about once per page, 274 snippets over 281 pages) to parts of the \coq development, showing definitions (Fig.~\ref{fig:Hprime}), computation results (Fig.~\ref{fig:Pow-evalPow17LetIn}), lemmas (Fig.~\ref{fig:Ex42E0}) and parts of  proof scripts (Fig.~\ref{fig:Ex42-schutte}). The order in which these references appear in the book is independent of the structure of our libraries. Thus, we chose not to include our textbook as comments in the libraries (solution \emph{à la} Coqdoc).
%\TODO[Pierre]{is the preceding sentence OK?}

@@

   \begin{figure}[h]
  \centering
  \inputsnippets{Pow/evalPow17LetIn}
  \caption{\label{fig:Pow-evalPow17LetIn}Automatically capturing the output of computations}
\end{figure}

@@

@@

The built artifacts are stored in the \community binary cache on Cachix~\cite{cachix}.
%
This means that a given version never has to be built twice and Continuous Integration can be almost instantaneous when no change has been made (e.g., after merging a pull request).
